---
title: "General Social Survey"
author: "Claudia Hernandez"
date: "Summer 2024"
output: 
  html_document:
    number_sections: true
    toc: true
---


```{r include=FALSE,echo=FALSE}
require(tigerstats)
require(tidyverse)
require(foreign)
source("GrabData.R")
```




#  Question 1

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->


The data in the dataframe GSSdata is from the 2018 General Social Survey. The first blocks of R-code has selected down a subset of the data to just 16 variables. It has further removed unwanted factor levels in much of the data. Examine the code in the GrabData.R file to see what it is doing. Some of the variables are categorical and others are numerical. Be sure to do a variable analysis before tackling each question.  
First question - Is opinion on the death penalty (CAPPUN) independent of gun ownership (OWNGUN)?

$H_0$: CAPPUN is independent of aown a gun, they are not related.  
$H_A$: they are dependent to owns the gun. 

## Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->

CAPPUN is a carigirical variable,  
OWNGUN is a carigorical variable, so I   
use the cat-cat helper file   
We want CPPUN~OWNGUN  
##Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

### Descriptive Results

#### Graphical Descriptive Results
```{r}
barchartGC(~OWNGUN + CAPPUN,data=GSSdata)
barchartGC(~OWNGUN +CAPPUN,data=GSSdata, type="percent")

```

This shows that people who own a gun are likely to favor capital punishment.


<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->

#### Numerical Descriptive Results
```{r}
table1 <- xtabs(~ OWNGUN + CAPPUN,data=GSSdata)
rowPerc(table1)
colPerc(table1)
```
This shows that who wons a gun 75% is favor and 57% is opposed. 


<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

### Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The p-value for the Chi- squared test was p = 0.02022
the mean we reject /fail to reject the null hpothesis , and 
Indepentent of gun ownership (OWNGUN) because its less than .05. 

The fisherexact test has a p-value of .01651 hence we rject
the null hypothesis,as with chi-square test, and furthermore the odds ratio says that gun owners are 2.3 times more likely to favor capital punishment than thos who do not own a gun.

#  Question 2

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

Is there a difference in years of education (EDUC) between the those who did or did not vote in 2012 (VOTE12)?

$H_0$: There is a difference in years in education between thoes who did and did not vote.   
$H_A$: There is a difference. 


## Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->

The EDUC is a numerical variable   
The voting is a categorical variable   
I used the numerical - cat.rmd file to pull the data.   

##Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

### Descriptive Results

#### Graphical Descriptive Results
```{r}
plot(EDUC~VOTE12,data=GSSdata)
```
   This shows that more people who voted in 2012 tend to have more education.
 
<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->

#### Numerical Descriptive Results


```{r}
favstats(EDUC~VOTE12,data=GSSdata)
```

This chart demonstrates that individuals that voted12 has an average of 16 years of education and people who did not vote had 13 years of education.

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

### Inferential Results
```{r}
t.test(EDUC~VOTE12,data=GSSdata)
```
The p-value is 0.00000001 so this means we reject the null hypothesis of because the p-value is smaller

I conclude that there is a difference in years of education between those who did or did not vote in 2012. 

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

#  Question 3

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

Determine if a persons performance on the wordsum test (WORDSUM) is independent of their level of schooling (DEGREE).
$H_0$:The score of a persons wordsum test is independent of there level of schoolong. 
$H_A$: They are dependent 

## Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->
wordsum is numerical variable   
Degree is categorical variable    
I used the NUM~CAT data file.   

##Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

### Descriptive Results

#### Graphical Descriptive Results


```{r}
boxplot(WORDSUM~DEGREE,data=GSSdata)
```
This graph shows that there is a difference between the performance of a persons testing and the level of schooling. 


<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->

#### Numerical Descriptive Results

```{r}
favstats(WORDSUM~DEGREE,data=GSSdata)
```
The graph shows there is a difference between LT high school and gradates because LT High is at at 7 and graduate median is at 10. 

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

### Inferential Results

```{r}
anova=aov(WORDSUM~DEGREE,data=GSSdata)
summary(anova)
```

This shows that we reject the null because the p-value is 0.0000002 witch is smaller then .05


```{r}
#we specify the margins starting at the bottom and proceeding clockwise - thus 8.7 is the left side margin to hold the names of the factors
par(mar = c(7, 8.7, 2, 2))
plot(TukeyHSD(anova),las=2)
```
There is no difference bwteen high school and less than high  
There is a difference and jr college and high schhol   
There is a difference between bachelors and high school   
There is a difference and gradutes and high school   
There is no difference betwwen high school and gradute   
No diffidence between jr college and grduate   
No difference between and bachelors and gradute   

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->


# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->


#### Question 1:is gun owerships independent of opinion on capital punishment. 
The summary results and the inferential results show that they ae not independent.

#### Question2: Is there a difference in years of education (EDUC) between the those who did or did not vote in 2012 (VOTE12)?
In graphical results it shows that more people who voted in 2012 tend to have more education

#### Question 3: Determine if a persons performance on the wordsum test (WORDSUM) is independent of their level of schooling (DEGREE).
We determined in inferential results that there is a diffrence between the performance on the WORDSUM test report. 
